---
title: "baseline"
author: "Ethan de Villiers"
date: "`r Sys.Date()`"
output: html_document
---

## Init
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r packages, include=FALSE}

library(devtools) # Essential for Aurum integration
library(tidyverse)  # streamlining R coding since 2016
library(r2r) # Used in Hashmaps later
library(RMySQL)

install_github('drkgyoung/Exeter_Diabetes_aurum_package') # Package created by Exeter University to work with CPRD data
library(aurum)
```

``` {r cprd, include=FALSE, echo=FALSE}
cprd = CPRDData$new(cprdEnv = "test-remote", cprdConf = '~/RStudio/aurum.yaml')
```

## Pulling Biomarkers
biomarkers <- c("weight", "height", "bmi", "fastingglucose", "hdl", "triglyceride", "creatinine_blood", "ldl", "alt", "ast", "totalcholesterol", "dbp", "sbp", "acr", "albumin_blood", "bilirubin", "haematocrit", "haemoglobin", "pcr")
 - Keep HbA1c separate as processed differently
 
Compelted Biomarkers:
 - Weight
 - Height
 - BMI
 - 

``` {r Pulling Baseline BMI}
analysis = cprd$analysis("all")

bmi_data = bmi_data %>% 
  analysis$cached("patid_clean_bmi_medcodes")
t1t2_cohort = t1t2_cohort %>%
  analysis$cached("t1t2_cohort")

analysis = cprd$analysis("ethan")
raw_bmi = t1t2_cohort %>%
  select(patid, dm_diag_date_all) %>%
  left_join(bmi_data, by="patid") %>%
  mutate(date_diff = datediff(date, dm_diag_date_all)/7) %>%
  filter(date_diff >= -12 & date_diff <=6) %>%
  group_by(patid) %>%
  mutate(min_date_diff=min(abs(date_diff), na.rm=TRUE)) %>%
  dbplyr::window_order(date_diff) %>%
  filter(row_number()==1) %>%
  ungroup() %>%
  filter(abs(date_diff)==min_date_diff) %>%
  analysis$cached("atdiagnosis_raw_bmi")

t1t2_cohort = t1t2_cohort %>%
  left_join(
    raw_bmi %>% 
    select(patid, bmi=testvalue, bmi_date=date),
    by="patid"
  ) %>%
  analysis$cached("atdiagnosis_bmi_clean") # change to atdiagnosis_v1_bmi

```

``` {r Calculating Time to Insulin}
# Katie has already found earliest insulin date and diagnosis date: 'dm_diag_insdate' and 'dm_diag_date_all'
t1t2_cohort = t1t2_cohort %>%
  mutate(time_to_insulin_weeks = datediff(dm_diag_insdate, dm_diag_date_all)/7) %>%
  analysis$cached("atdiagnosis_v2_instime")

```

``` {r Pulling Baseline Weight}
analysis = cprd$analysis("all")

# Obtaining new Weight Table to pull into my data
weight_data = weight_data %>% 
  analysis$cached("patid_clean_weight_medcodes")

# Obtaining my table to add to
analysis = cprd$analysis("ethan")
my_cohort = my_cohort %>%
  analysis$cached("atdiagnosis_v2_instime")

raw_weight = my_cohort %>%
  select(patid, dm_diag_date_all) %>%
  left_join(weight_data, by="patid") %>%
  mutate(date_diff = datediff(date, dm_diag_date_all)/7) %>%
  filter(date_diff >= -12 & date_diff <=6) %>%
  group_by(patid) %>%
  mutate(min_date_diff=min(abs(date_diff), na.rm=TRUE)) %>%
  dbplyr::window_order(date_diff) %>%
  filter(row_number()==1) %>%
  ungroup() %>%
  filter(abs(date_diff)==min_date_diff) %>%
  analysis$cached("atdiagnosis_raw_weight")

my_cohort = my_cohort %>%
  left_join(
    raw_weight %>% 
    select(patid, weight=testvalue, weight_date=date),
    by="patid"
  ) %>%
  analysis$cached("atdiagnosis_v3_weight")

```

``` {r Pulling Baseline Height}

analysis = cprd$analysis("all")

# Obtaining new Weight Table to pull into my data
height_data = height_data %>% 
  analysis$cached("patid_clean_height_medcodes")

# No need to assign my_cohort as it was assigned to _v3_weight in last section
analysis = cprd$analysis("ethan")

raw_height = my_cohort %>% # takes about 10 mins to run on Slade
  select(patid, dm_diag_date_all) %>%
  left_join(height_data, by="patid") %>%
  mutate(date_diff = datediff(date, dm_diag_date_all)/7) %>%
  filter(date_diff >= -12 & date_diff <=6) %>%
  group_by(patid) %>%
  mutate(min_date_diff=min(abs(date_diff), na.rm=TRUE)) %>%
  dbplyr::window_order(date_diff) %>%
  filter(row_number()==1) %>%
  ungroup() %>%
  filter(abs(date_diff)==min_date_diff) %>%
  analysis$cached("atdiagnosis_raw_height")

my_cohort = my_cohort %>%
  left_join(
    raw_height %>% 
    select(patid, height=testvalue, height_date=date),
    by="patid"
  ) %>%
  analysis$cached("atdiagnosis_v4_height")

```

``` {r Pulling Baseline Hba1c}

analysis = cprd$analysis("all")

# Obtaining new Weight Table to pull into my data
hba1c_data = hba1c_data %>% 
  analysis$cached("patid_clean_hba1c_medcodes")

# So... Always re-assign to the latest version to avoid versioning issues :(
analysis = cprd$analysis("ethan")

my_cohort = my_cohort %>%
  analysis$cached("atdiagnosis_v4_height")

raw_hba1c = my_cohort %>% # takes about 10 mins to run on Slade
  select(patid, dm_diag_date_all) %>%
  left_join(hba1c_data, by="patid") %>%
  mutate(date_diff = datediff(date, dm_diag_date_all)/7) %>%
  filter(date_diff >= -12 & date_diff <=6) %>%
  group_by(patid) %>%
  mutate(min_date_diff=min(abs(date_diff), na.rm=TRUE)) %>%
  dbplyr::window_order(date_diff) %>%
  filter(row_number()==1) %>%
  ungroup() %>%
  filter(abs(date_diff)==min_date_diff) %>%
  analysis$cached("atdiagnosis_raw_hba1c")

my_cohort = my_cohort %>%
  left_join(
    raw_hba1c %>% 
    select(patid, hba1c=testvalue, hba1c_date=date),
    by="patid"
  ) %>%
  analysis$cached("atdiagnosis_v5_hba1c")

```

## Calculations Section

``` {r Calculating Model 1: Clinical Features}
### Model 1 logOddsRatio = 37.94 + (-5.09 * log(dm_diag_age_all)) + (-6.34 * log(bmi))
### Model Probability = exp(logOR) / (1 + exp(logOR))

# model_data = model_data %>%
#   mutate(Model5Prob = ifelse(!is.na(Model5LogOR), exp(Model5LogOR) / (1 + exp(Model5LogOR)), NA))

#Update this as I continue pulling baseline characteristics
my_cohort = my_cohort %>%
  analysis$cached("atdiagnosis_v5_hba1c")

my_cohort = my_cohort %>%
  mutate(dm_diag_age_all = ifelse(dm_diag_age_all < 0, 0, dm_diag_age_all)) %>%
  mutate(model_clinfeats = ifelse(!is.na(dm_diag_age_all) & !is.na(bmi) & dm_diag_age_all != 0,
                               exp(37.94 + (-5.09 * log(dm_diag_age_all)) + (-6.34 * log(bmi))) / (1 + exp(37.94 + (-5.09 * log(dm_diag_age_all)) + (-6.34 * log(bmi)))),
                               NULL)) %>%
  analysis$cached("atdiagnosis_vX_models")


```


``` {r Legacy Code}
#Pulling data
analysis = cprd$analysis("all")
t1t2_cohort = t1t2_cohort %>% 
  analysis$cached("t1t2_cohort") %>%
  mutate(BMI = NA) %>%
  analysis$cached('ethan_test')

t1t2_cohort_test = t1t2_cohort_test %>%
  analysis$cached("t1t2_cohort")

clean_hba1c_clean_units = clean_hba1c_clean_units %>%
  analysis$cached("patid_clean_hba1c_medcodes")


View(t1t2_cohort)

# Pull BMI data
analysis = cprd$analysis("all_patid_clean")
bmi_data = bmi_data %>% 
  analysis$cached("bmi_medcodes") %>%
  cbind.data.frame()

#Cleaning/preparing datasets
bmi_data = bmi_data %>%
  mutate(patid = as.character(patid)) # Change to strings to ensure large int integrity
t1t2_cohort = t1t2_cohort %>%
  mutate(patid = as.character(patid)) # Change to strings to ensure large int integrity

t1t2_cohort$bmi_date = NA
t1t2_cohort$bmi = NA

# Obtain all Patient IDs from T1T2 cohort that we need to calculate baseline BMI
patid_list = t1t2_cohort$patid

# Obtain all Patient IDs from T1T2 cohort that we need to calculate baseline BMI
bmi_patid = patid_list

for (i in 1:length(bmi_patid)) { 
  date_diag = t1t2_cohort %>%
    filter(bmi_patid[i] == patid) %>%
    select(dm_diag_date_all) 
  
  date_diag = date_diag$dm_diag_date_all[1] # Extract pure value of date_diag
  print(paste('NEW Patient! Diag Date:', date_diag)) # Now have patient's diagnosis date
  
  # Now need to convert patient's BMI scores to a list
  bmi_list = bmi_data %>%
    filter(patid == bmi_patid[i]) %>%
    select(date)
  
  bmi_list = as.Date(bmi_list$date, format="%Y-%m-%d")
  
  # Skip over this iteration of loop if no BMI data
  if (length(bmi_list) == 0){
    print(paste('No BMI results for patient: ', patid_list[i]))
    patid_list = patid_list[patid_list != bmi_patid[i]]
    next 
  } 
  
  # Logging
  print(paste('number of BMI entries = ', length(bmi_list)))
  print(bmi_list)
  
  dates_and_differences = hashmap()
  
  for (x in 1:length(bmi_list)){
    individ_date_diff = difftime(bmi_list[x], date_diag, units="weeks")
    if (individ_date_diff >= -12 & individ_date_diff <= 6){
      print(paste('date is accepted! >-12 & <6, diff = ', individ_date_diff))
      # Date is more recent than 12 weeks before diagnosis, and more recent than 6 weeks after diagnosis
      #insert(dates_and_differences, key = )
      dates_and_differences[[bmi_list[x]]] = abs( individ_date_diff ) # Saves absolute difference
    }
  }
  
  
  # SIf patients does not have any valid BMI readings 12 weeks before or 6 weeks after diagnosis, skip iteration
  if (length(dates_and_differences) == 0){
    print('No acceptable dates! Skipping patient...')
    patid_list = patid_list[patid_list != bmi_patid[i]]
    next 
  } 
  
  
  dates = as.numeric(values(dates_and_differences)) # convert difftie object into DOUBLE type
  dates = sort(dates) # Order absolute values in ascending order
  
  print(paste('closest to diag date = ', dates[1], ' weeks difference!'))
  
  for (x in 1:length(dates_and_differences)){
    key = keys(dates_and_differences)[[x]]
    if (dates_and_differences[[key]] == dates[1]){
      print(paste('BMI Date match found in Hashmap. Date stored = ', as.character(key)))
      bmi_diag_date = as.character(key)
    }
  }
  
  bmi_value = bmi_data %>%
    filter(bmi_patid[i] == patid) %>%
    filter(date == bmi_diag_date)
  bmi_value = bmi_value$testvalue
  print(paste('Pulled bmi_value from patient according to hashmap date stored. BMI pulled = ', bmi_value))
  
  #Now have the correct diagnosis date: bmi_diag_date = BMI diagnosis date
  #And have the correct diagnosis value: bmi_value = BMI test result on said date
  t1t2_cohort = t1t2_cohort %>%
    mutate(bmi_date = ifelse(bmi_patid[i] == patid, bmi_diag_date, bmi_date)) %>%
    mutate(bmi = ifelse(bmi_patid[i] == patid, bmi_value, bmi))
  
  

  # Remove patID from original list without interrupting this for loop iteration as it BMI has been completed for this patient!
  patid_list = patid_list[patid_list != bmi_patid[i]]
}



#db.host = readline(prompt="Enter 'localhost' or Server IP:")
db.username = readline(prompt="Enter MySQL Username: ")
db.password = readline(prompt="Enter MySQL Password: ")
db.database = readline(prompt="Enter Target Database: ")

# Connect to a MySQL database running locally
con <- dbConnect(RMySQL::MySQL(), dbname = db.database, user = db.username, password = db.password) # host = db.host

dbWriteTable(conn = con, name = 'ethan_test_dbconnection', value = t1t2_cohort)

# Always cleanup by disconnecting the database
dbDisconnect(con)
```