---
title: "Thyroid investigation"
author: "Ethan de Villiers"
date: "`r Sys.Date()`"
output: html_document
---

``` {r Init, include = FALSE} 

library(devtools) # Essential for Aurum integration
library(tidyverse)  # streamlining R coding since 2016
library(r2r) # Used in Hashmaps later
library(RMySQL)

install_github('drkgyoung/Exeter_Diabetes_aurum_package') # Package created by Exeter University to work with CPRD data
library(aurum)

cprd = CPRDData$new(cprdEnv = "test-remote", cprdConf = '~/RStudio/aurum.yaml')
codesets = cprd$codesets()
codes = codesets$getAllCodeSetVersion(v = "31/10/2021")
```

# Looking at TTG and TSH within Diabetic children (<18yo)

``` {r Descriptive stats}
analysis = cprd$analysis("all")

t1t2 = t1t2 %>%
  analysis$cached("t1t2_cohort")

thyroid_cohort = thyroid_cohort %>%
  left_join()

observation_ttg <- cprd$tables$observation %>%
  filter(medcodeid == codes$)


# Delete
https://github.com/Exeter-Diabetes/All-Github-Guides/blob/main/1-Nomenclature.md
https://github.com/Exeter-Diabetes/All-Github-Guides/blob/main/2-GitConfig.md
https://github.com/Exeter-Diabetes/All-Github-Guides/blob/main/3-Repositories.md
https://github.com/Exeter-Diabetes/All-Github-Guides/blob/main/4-GitBash.md
https://github.com/Exeter-Diabetes/All-Github-Guides/blob/main/5-RStudioGit.md

Before we can start working on the Github repository, we have to _clone_ the repository into our local environment. This means creating a new project folder on our computer, which will hold all of the files locally. We then initialise **Git** in the folder, so that we have the Git tools required to connect to our online repository! Once its cloned (one of the Git tools), we can open this folder in our IDE (such as R-Studio) and begin adding files, creating scripts, or writing code.


  


```