---
title: "Thyroid investigation"
author: "Ethan de Villiers"
date: "`r Sys.Date()`"
output: html_document
---

``` {r Init, echo=FALSE} 

library(devtools) # Essential for Aurum integration
library(tidyverse)  # streamlining R coding since 2016

install_github("Exeter-Diabetes/CPRD-analysis-package") # Package created by Exeter University to work with CPRD data
library(aurum)
cprd = CPRDData$new(cprdEnv = "test-remote", cprdConf = '~/RStudio/aurum.yaml')
codesets = cprd$codesets()
codes = codesets$getAllCodeSetVersion(v = "31/10/2021")
```

# Looking at TTG and TSH within Diabetic children (<18yo)

``` {r Creating raw_tsh table for ThyroidInv, echo=FALSE}
#Obtain TSH data
analysis = cprd$analysis("all")
thyroid_data = thyroid_data %>% 
  analysis$cached("patid_clean_tsh")

#Obtain t1t2_cohort for list of patid (relevant patients)
t1t2_cohort = t1t2_cohort %>%
  analysis$cached("t1t2_cohort")

analysis = cprd$analysis("ethan")
raw_tsh = t1t2_cohort %>%
  select(patid, dm_diag_date_all) %>%
  left_join(thyroid_data, by="patid") %>%
  mutate(date_diff = datediff(obsdate, dm_diag_date_all)/7) %>%
  filter(date_diff <= -4) %>%
  group_by(patid) %>%
  mutate(min_date_diff=min(abs(date_diff), na.rm=TRUE)) %>%
  dbplyr::window_order(date_diff) %>%
  filter(row_number()==1) %>%
  ungroup() %>%
  filter(abs(date_diff)==min_date_diff) %>%
  analysis$cached("thyroidinv_raw_tsh")


remove(thyroid_data)
```

``` {r Creating raw_ttg table for ThyroidInv, echo=FALSE}
#Obtain TSH data
analysis = cprd$analysis("all")
thyroid_data = thyroid_data %>% 
  analysis$cached("patid_clean_ttg")

#Obtain t1t2_cohort for list of patid (relevant patients)
t1t2_cohort = t1t2_cohort %>%
  analysis$cached("t1t2_cohort")

analysis = cprd$analysis("ethan")
raw_ttg = t1t2_cohort %>%
  select(patid, dm_diag_date_all) %>%
  left_join(thyroid_data, by="patid") %>%
  mutate(date_diff = datediff(obsdate, dm_diag_date_all)/7) %>%
  filter(date_diff <= -4) %>%
  group_by(patid) %>%
  mutate(min_date_diff=min(abs(date_diff), na.rm=TRUE)) %>%
  dbplyr::window_order(date_diff) %>%
  filter(row_number()==1) %>%
  ungroup() %>%
  filter(abs(date_diff)==min_date_diff) %>%
  analysis$cached("thyroidinv_raw_ttg")



```

``` {r Creating final thyroid_cohort}
# merge both cohorts onto V7 for antibody + bmi + other data
analysis = cprd$analysis("ethan")
v7_cohort = v7_cohort %>%
  analysis$cached("atdiagnosis_v7_ia2")

thyroid_cohort = v7_cohort %>%
  left_join(
    raw_tsh %>% 
    select(patid, tsh=result, tsh_date=obsdate),
    by="patid"
  ) %>%
  left_join(
    raw_ttg %>% 
    select(patid, ttg=result, ttg_date=obsdate),
    by="patid"
  ) %>%
  analysis$cached("thyroidinv_cohort")
```

Final cohort is now discoverable under the MySQL table as "ethan_thyroidinv_cohort"
